---
title: "StatisticalAnalysis_AFQ"
output:
  html_document: default
editor_options:
  chunk_output_type: inline
---

setwd("/Users/YQ/OneDrive/Research/Shenzhen/2.ASD/4.ASDDTI")
here::set_here(path = "/Users/YQ/OneDrive/Research/Shenzhen/2.ASD/4.ASDDTI")

detach("package:here", unload=TRUE)

# Set up 
```{r message=FALSE, results='hide'}
packages <- c("here","psych","dplyr","tidyr", "readr", "broom", "ggplot2","perm",
              "gridExtra","ggpubr","car","purrr","stringr")

invisible(lapply(packages, library, character.only = TRUE))
```

# Prepare data
```{r}
AFQ_data <- read_csv("data/merged.csv",show_col_types = FALSE)
Beh <- readxl::read_excel("data/DTI_info_sorted_8.13.xlsx", sheet = 1)
colnames(Beh)
print(unique(AFQ_data$tractID))
AFQ_data <- AFQ_data %>%
  filter(tractID %in% c("Left_Arcuate","Right_Arcuate",
                        "Left_Inferior_Longitudinal","Right_Inferior_Longitudinal",
                        "Left_Superior_Longitudinal","Right_Superior_Longitudinal",
                        "Left_Inferior_Fronto_occipital",
                        "Right_Inferior_Fronto_occipital"))
print(colnames(Beh))
colnames(Beh)[12:14] <- c("ADOS_SA","ADOS_RRB","ADOS_Total")
Beh <- within(Beh, {
  Dx <- as.factor(Dx)
  Gender <- as.factor(Gender)
  Age <- as.numeric(Age)
  ADOS_SA <- as.numeric(ADOS_SA)
  ADOS_RRB <- as.numeric(ADOS_RRB)
  ADOS_Total <- as.numeric(ADOS_Total)
  CARS <- as.numeric(CARS)
})

tmp <- merge(AFQ_data,Beh,by = 'subject_id' )
colnames(tmp)[4:7]
data <- tmp %>% 
  pivot_longer(cols = colnames(tmp)[4:7], names_to = "metric", 
               values_to = "value")

# Calculate the average DTI metrics across the 100 nodes for each subject
# within each tract
data_avg <- data %>%
  group_by(subject_id, tractID, metric) %>%
  summarise(
    avg_value = mean(value),
    across(c(Gender, Age, Dx, Gesell_Adpt, Gesell_GrM, Gesell_FineM, Gesell_Lang, 
             Gesell_Social, Gesell_Total, ADOS_SA, ADOS_RRB, ADOS_Total), first),
    .groups = 'drop'
  )

View(data_avg)
write.csv(data_avg, 'data/FAQ_Beh_data.csv') 
# write.csv(data_avg, '../results/all_data_avg.csv') 

```

# Descriptive statistics: 
```{r}
Beh$Dx <- factor(Beh$Dx, levels = c("TD", "ASD"))

colnames(Beh)[c(5,4,16,15,12:14,6:11)]
mean_all <- describeBy(Beh[,c(5,4,16,15,12:14,6:11)], group ="Dx")


demo <- as.data.frame(
  cbind(paste0(round(mean_all[[1]][2:13,]$mean,2), "±",
               round(mean_all[[1]][2:13,]$sd,2)),
        paste0(round(mean_all[[1]][2:13,]$min,2), "-",
               round(mean_all[[1]][2:13,]$max,2)),
        paste0(round(mean_all[[2]][2:13,]$mean,2), "±",
               round(mean_all[[2]][2:13,]$sd,2)),
        paste0(round(mean_all[[2]][2:13,]$min,2), "-",
               round(mean_all[[2]][2:13,]$max,2))
  ))


rownames(demo) <- rownames(mean_all[[1]])[2:13]

for (bb in rownames(demo)[c(1,2,7:12)]){
  
  if (!all(is.na(Beh[Beh$Dx == "TD", bb]))) {
    t_test <- t.test(Beh[Beh$Dx=="TD",bb],Beh[Beh$Dx=="ASD",bb],
                      var.equal = TRUE,na.action = T)
    demo[bb,5:6] <- c(round(t_test$statistic,3),round(t_test$p.value,3))
  }
}

# Gender
df <- table(Beh$Gender, Beh$Dx)
demo[13,c(1,3)] <- c(paste0(df[2,1],"/",df[1,1]),
                     paste0(df[2,2],"/",df[1,2]))
gender_fisher <- fisher.test(df)
demo[13,c(5:6)] <- round(c(gender_fisher$estimate,gender_fisher$p.value),2)
rownames(demo)[13] <- "Gender(M/F)"
colnames(demo) <- c("TD_mean±SD","TD_range","ASD_mean±SD","ASD_range",
                    "Estimate","p_value")

demo <- demo[c(1,13,2:12),]
demo[4:7,1:2] <- NA

# write.csv(demo,'../results/Descriptive.csv')
```

# Lateralization analysis and barplots
```{r}
## prepare data for lateralization analysis
roi <- c("Left_Arcuate","Left_Inferior_Fronto_occipital",
         "Right_Arcuate","Right_Inferior_Fronto_occipital",
         "Left_Inferior_Longitudinal","Left_Superior_Longitudinal",
         "Right_Inferior_Longitudinal","Right_Superior_Longitudinal")

data_new <- data_avg %>% filter(tractID %in% roi)

#View(data_new)

data_new$tractName <- sub("^(Left_|Right_)", "", data_new$tractID)
data_new$hem <- sub("_(.*)$", "", data_new$tractID)

data_lat <- data_new %>% 
  group_by(subject_id, tractName, metric) %>%
  reframe(lat = (avg_value[hem=="Right"]-avg_value[hem=="Left"])/(avg_value[hem=="Right"]+avg_value[hem=="Left"]),
            Dx = first(Dx),
            Age = first(Age),
            Gender = first(Gender),
            Gesell_Lang = first(Gesell_Lang))

write.csv(data_lat, "results/Lateralization_Index.csv", row.names = FALSE)

##  One-sample t-test
set.seed(123)

data_lat <- read.csv("results/Lateralization_Index.csv")
lateralization_one_sample_ttest <- data.frame()
i <- 0
for (dd in unique(data_lat$metric)){
  for (tt in unique(data_lat$tractName)){
    for (gg in c("ASD","TD")){
    i <- i+1
    df_new <- data_lat %>% filter(metric == dd & tractName == tt &
                              Dx == gg)
    
    # one-sample t-test
    observed_t <- as.numeric(t.test(df_new$lat, mu = 0)$statistic)
    observed_p <- as.numeric(t.test(df_new$lat, mu = 0)$p.value)
    
    # permutation test
    perm_t_values <- numeric()
    for (j in 1:5000) {
      perm_data <- df_new
      n <- length(perm_data$lat)
      centered_x <- perm_data$lat - mean(perm_data$lat)
      
      perm_sample <- sample(centered_x, n, replace = TRUE)
 
      perm_t_values[j] <- as.numeric(t.test(perm_sample, mu = 0)$statistic)
    }
    
    p_value_perm <- mean(abs(perm_t_values) >= abs(observed_t))
    
    
    mean_value <-round(mean(df_new$lat),3)
    sd_value <- round(sd(df_new$lat),3)
    se_value <- round(sd_value/sqrt(nrow(df_new)),3)
    
    lateralization_one_sample_ttest[i,1:9] <- c(dd,tt,gg,
                                                mean_value,sd_value,se_value,
                                                observed_t,
                                                observed_p,
                                                p_value_perm)
    
    }
  }
}
colnames(lateralization_one_sample_ttest) <- c("metric","tractName","Dx",
                                               "mean","sd","se","t_value",
                                               "p_value","p_value_perm")
write.csv(lateralization_one_sample_ttest,
          "results/LI_one_sample_ttest_with_permutation.csv", 
          row.names = FALSE)

##  two-sample t-tests
set.seed(123)
data_lat <- read.csv("results/Lateralization_Index.csv")

lateralization_two_sample_ttest <- data.frame()
i <- 0
for (dd in unique(data_lat$metric)){
  for (tt in unique(data_lat$tractName)){
      
    i <- i+1
    df_new <- data_lat %>% filter(metric == dd & tractName == tt )
    
    # two-sample t-test
    t_test_result <- t.test(lat ~ Dx, data=df_new)
    observed_t <- t_test_result$statistic
    observed_p <- t_test_result$p.value
    
    # permutation test
    perm_t_values <- numeric()
    for (j in 1:5000) {
      perm_data <- df_new
      perm_data$Dx <- sample(perm_data$Dx)
      perm_t_values[j] <- t.test(lat ~ Dx, data=perm_data)$statistic
    }
    p_value_perm <- mean(abs(perm_t_values) >= abs(observed_t))
    
    lateralization_two_sample_ttest[i,1:5] <- c(dd,tt,observed_t,observed_p,
                                                p_value_perm)
  }
}
colnames(lateralization_two_sample_ttest) <- c("metric","tractName",
                                               "t_value","p_value","p_value_perm")

write.csv(lateralization_two_sample_ttest,
          "results/LI_two_sample_ttest_with_permutation.csv", row.names = FALSE)


## bar plots for LI
aa <- read.csv("results/LI_one_sample_ttest_with_permutation.csv")
View(aa)
for (dd in unique(aa$metric)){
  df <- aa %>% filter(metric == dd)
  p <- ggplot(df, aes(x = roi, y = mean, fill = Dx)) +
    geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se),size=1.5,
                  width = 0.2, position = position_dodge(0.7)) +
    geom_hline(yintercept = 0, color = "black",size=1) + 
    labs(title = paste0(dd), 
           x = NULL, y = NULL) +
    theme_classic() +
    theme(
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5), 
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.length = unit(0.25, "cm"), 
      axis.ticks.x = element_line(size=1), 
      axis.ticks.y = element_line(size=1), 
      axis.line.x = element_line(size = 2.5),
      axis.line.y = element_line(size = 2.5),
      legend.position = "none")+
    scale_fill_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC"))
  
    ggsave(paste0("results/",dd,".png"),plot = p, width = 12, height = 10)
}

```

# Correlation between LI and Gesell language
```{r}
set.seed(123)
correlation_results <- data.frame()

data_lat <- read.csv("results/Lateralization_Index.csv")

tmp <- data_lat
tmp$Dx <- "All"
df <- rbind(data_lat, tmp)

df$gender[df$Gender == "M"] <- 1
df$gender[df$Gender == "F"] <- 2

i <- 0
for (dd in unique(df$metric)){
  for (tt in unique(df$tractName)){
    for (gg in unique(df$Dx)){
      i <- i+1
      df_new <- df %>% filter(metric == dd & tractName == tt &
                                        Dx == gg)
  
      formula <- formula(lat ~ Gesell_Lang + Age + Gender)
      model <- lm(formula, data = df_new)
      
      observed_coeff <- summary(model)$coefficients[2, "Estimate"]
      observed_p_value <- summary(model)$coefficients[2, "Pr(>|t|)"]
      
      
      cor_r <- cor.test(df_new$lat, df_new$Gesell_Lang)
      
      ppcor <- ppcor::pcor.test(df_new$lat, df_new$Gesell_Lang,
      			  df_new[,c("Age","gender")])

      
      # Perform permutation test
      #permuted_coeffs <- numeric()
      #for (j in 1:5000) {
      #  perm_data <- df_new
      #  perm_data$lat <- sample(perm_data$lat)
       # permuted_model <- lm(lat ~ Gesell_Lang + Age + Gender, data = perm_data)
       # permuted_coeffs[j] <- summary(permuted_model)$coefficients[2, "Estimate"]
      # }
      
      #p_value_perm <- mean(abs(permuted_coeffs) >= abs(observed_coeff))
      
      correlation_results[i,1:7] <- c(dd,tt,gg,observed_coeff,observed_p_value,
                                     cor_r$estimate,ppcor$estimate)
    
    }
  }
}

colnames(correlation_results) <- c("DTI_metric","tractName","Dx",
                                   "regression_coefficients",
				   "P_value","r_value","ppr_value")

View(correlation_results)

write.csv(correlation_results, "results/cor_LI_Lang_v2.csv", row.names = FALSE)

## significant correlation results
sig_cor <- correlation_results[correlation_results$perm_p_value < 0.05,]
correlation_results[correlation_results$DTI_metric == sig_cor$DTI_metric[] & 
		    	correlation_results$tractName == sig_cor$tractName[2],][1:2,]

## scatter plots
aa <- read.csv("results/cor_LI_Lang.csv")
data_lat <- read.csv("results/Lateralization_Index.csv")

df <- aa %>% filter(P_value < 0.05)

df$DTI_metric <- gsub("lateralization", "avg_dti", df$DTI_metric)

for (i in 1:nrow(df)){
  data_new <- data_lat %>% filter(metric == df$DTI_metric[i] &
                                    tractName == df$tractName[i])
  
  formula <- formula(lat ~ Gesell_Lang + Age + Gender)
  model_ASD <- lm(formula, data = data_new[data_new$Dx == "ASD",])
  model_TD <- lm(formula, data = data_new[data_new$Dx == "TD",])
  pred_ASD <- predict(model_ASD, newdata = data_new[data_new$Dx == "ASD",])
  pred_TD <- predict(model_TD, newdata = data_new[data_new$Dx == "TD",])
  data_new$pred[data_new$Dx == "TD"] <- pred_TD
  data_new$pred[data_new$Dx == "ASD"] <- pred_ASD
  
  title <- paste0(df$DTI_metric[i],"_",df$tractName[i])
  
  p <- ggplot(data_new, aes(x = Gesell_Lang, y = lat, color = Dx)) +
    geom_point(aes(color = Dx),size = 3.5) +
    stat_smooth( method = "lm", aes(y = pred), se = FALSE,
                size = 1.8) +
    labs(title = title,x = NULL, y = NULL) +
    theme_classic() +
    theme(
      plot.title = element_text(size = 20, face = "bold", hjust = 0.5), 
      #axis.text = element_blank(),
      axis.line.x = element_line(size = 1.5),
      axis.line.y = element_line(size = 1.5),
      legend.position = "none")+
    scale_color_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC"))  
  
  ggsave(paste0("results/",title,"_lat.png"), plot = p, width = 5, height = 4)
}

```

# Group comparison at the entire tract 
```{r}
ANCOVA_mean_node_results <- data.frame()
set.seed(123)
i <- 0
for (dd in unique(data_avg$metric)){
  for (tt in unique(data_avg$tractID)){
    
    i <- i+1
    data_new <- data_avg %>% filter(tractID == tt & metric == dd)
    
    # ANOVA for each metric
    fl <- formula(avg_value ~ Dx + Age + Gender)
    anova_gr <- car::Anova(lm(fl, data_new), type ="III")
    observed_f <- anova_gr$`F value`[2]
    observed_p <- anova_gr$`Pr(>F)`[2]
    
    # Perform permutation test by shuffling the group labels
    
    perm_f_values <- numeric()
    for (j in 1:5000) {
      perm_data <- data_new
      perm_data$Dx <- sample(perm_data$Dx)
      perm_anova <- car::Anova(lm(fl, perm_data), type="III")
      perm_f_values[j] <- perm_anova$`F value`[2]
      }
    p_value_perm <- mean(abs(perm_f_values) >= abs(observed_f))
    
    # Calculate mean DTI values for ASD and TD groups
    ASD_mean <- mean(data_new$avg_value[data_new$Dx == "ASD"], na.rm = TRUE)
    TD_mean <- mean(data_new$avg_value[data_new$Dx == "TD"], na.rm = TRUE)
    
    ANCOVA_mean_node_results[i,1:7] <- c(dd,tt, ASD_mean, TD_mean,
                                        observed_f,
                                        observed_p,
                                        p_value_perm)
    
  }
}

colnames(ANCOVA_mean_node_results) <- c("metric","tractID","mean_ASD","mean_TD",
                                        "F_value","p_value","p_value_perm")

write.csv(ANCOVA_mean_node_results,'results/tract_mean_node_anova_results.csv',
	  row.names = FALSE)


## Violin plots showing for group comparison between ASD and TD for each DTI metric (FA, MD, RD, AD) within each tract

# Loop over each DTI metric
for (mm in unique(data_avg$metric)) {
  
  data_new <- data_avg %>% filter(metric == mm)
  
  # Create the plot
  p <- ggplot(data_new, aes_string(x = "tractID", y = "avg_value", fill = "Dx")) +
    geom_violin(trim = FALSE, alpha = 1, width = 0.8, aes(fill = Dx),size=1.5) +
    geom_boxplot(aes(Dx = interaction(Dx,"tractID")), 
                 position = position_dodge(0.8),
                 color = "white", size=1.5,
                 width = 0.2, show.legend = FALSE, outlier.shape=NA) +
    geom_point(position = position_jitterdodge(0.2), size = 3,color = "black") +
    labs(x = NA, y = NA) +
    theme_classic() +
    theme(axis.ticks.length = unit(0.25, "cm"), 
          axis.text = element_blank(),
          axis.title = element_blank(),
          axis.line.x = element_line(size = 2),
          axis.line.y = element_line(size = 2),
          legend.position = "none") +
    scale_fill_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC")) 
  
  # Print the plot to ensure it looks correct
  print(p)
  
  # Save the plot with a unique filename
  ggsave(filename = paste0('../figures/Violin/grouped_violin_', mm, '.png'), 
         plot = p, height = 6, width = 20, dpi = 300)
}

```

# Group comparison at point-wise levels within each tractID
```{r}
set.seed(123)

tract_each_node_anova_results <- data.frame()

i <- 0
for (dd in unique(data$metric)){
  for (tt in unique(data$tractID)){
    
    data_new <- data %>% filter(tractID == tt & metric == dd)
    
    for (nn in 0:99){
      i <- i+1
      data_new_nn <- data_new %>% filter(nodeID == nn)
      
      # Define the models for each scalar metric
      model <- lm(as.formula(value ~ Dx + Age + Gender), data = data_new_nn)
      
      ## Perform permutation tests using custom function
      observed_anova <- car::Anova(model,type="III")
      observed_statistic <- observed_anova$`F value`[2]
      observed_p <- observed_anova$`Pr(>F)`[2]
      
      # Permutation loop
      perm_f_values <- numeric()
      for (j in 1:5000) {
        perm_data <- data_new_nn
        perm_data$Dx <- sample(perm_data$Dx)
        perm_model <- lm(as.formula(value ~ Dx + Age + Gender), 
                         data = perm_data)
        perm_anova <- car::Anova(perm_model,type="III")
        perm_f_values[j] <- perm_anova$`F value`[2]
      }
      
      p_value_perm <- mean(perm_f_values >= observed_statistic, na.rm = TRUE)
      
      # Calculate mean DTI values for ASD and TD groups
      ASD_mean <- mean(data_new_nn$value[data_new_nn$Dx == "ASD"], na.rm = TRUE)
      TD_mean <- mean(data_new_nn$value[data_new_nn$Dx == "TD"], na.rm = TRUE)
    
      tract_each_node_anova_results[i,1:8] <- c(dd,tt,nn,ASD_mean,TD_mean,
                                              observed_statistic,
                                              observed_p,
                                              p_value_perm)
    }  
  }
}

colnames(tract_each_node_anova_results) <- c("metric","tractID","nodeID","mean_ASD","mean_TD",
                                        "F_value","p_value","p_value_perm")

write.csv(tract_each_node_anova_results,
	  "results/tract_each_node_f_test_results.csv", row.names = FALSE)

## Curve plots: group comparsion point-wise
# new data
data_curve <- read.csv("../results/tract_each_node_f_test_results.csv")

# old data
# data_curve <- read.csv("../tables/group_comparison_point_wise/tract_each_node_f_test_results.csv")

scale_map <- list(
  dti_fa = list(breaks = seq(0.2, 0.7, by = 0.1), limits = c(0.2, 0.7)),
  dti_rd = list(breaks = seq(0.00048, 0.00088, by = 0.0001), limits = c(0.00048, 0.00088)),
  dti_ad = list(breaks = seq(0.0010, 0.0018, by = 0.0002), limits = c(0.0010, 0.0018)),
  dti_md = list(breaks = seq(0.0008, 0.0011, by = 0.0001), limits = c(0.0008, 0.0011))
)

for (dd in unique(data_curve$metric)){
  for (tt in unique(data_curve$tractID)){

    df <- data %>% filter(metric == dd & tractID == tt)

    summary_data <- df %>% group_by(nodeID, Dx) %>% 
      summarise(mean_value = mean(value,na.rm = TRUE),
                sd_value = sd(value, na.rm = TRUE),
                min_value = min(value),
                max_value = max(value),
                .groups = "drop")
    
    # find consecutive idx 
    data_curve_new <- data_curve %>% filter(metric == dd & tractID == tt)
    idx <- which(data_curve_new$p_value < 0.05)
    idx <- idx -1
    target <- list()
    if (length(idx) >= 3) {
      start <- 1
      for (i in 2:length(idx)) {
        if (idx[i] != idx[i - 1] + 1) {
          if (i - start >= 3) {
            target[[length(target) + 1]] <- idx[start:(i - 1)]
            }
          start <- i 
        }
        }
      if (start <= length(idx) - 2) {
        target[[length(target) + 1]] <- idx[start:length(idx)]
      }
      }

    if (length(target)>0){
    # prepare for plot
    y_range <- max(df$value) - min(df$value)
    y_padding <- y_range * 0.1  # 10% of the range as padding
    y_min <- min(df$value) - y_padding
    y_max <- max(df$value) + y_padding
    
    
    # plot
    p <- ggplot(data = summary_data, aes(x = nodeID, y = mean_value, color = Dx)) +
      geom_line(size = 3, alpha = 1) +
      # geom_ribbon(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value, fill = Dx),
      #             alpha = 0.2, color = NA) +
      scale_x_continuous(breaks = seq(0, 100, by = 20)) +
      scale_y_continuous(breaks = scale_map[[dd]]$breaks, 
                         limits = scale_map[[dd]]$limits,
                         labels = NULL)+
      scale_color_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC")) +
      scale_fill_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC")) +
      theme_classic() +
      labs(title = paste0(dd," ",tt), x = NULL, y = NULL) +
      theme(axis.ticks.length = unit(0.25, "cm"), 
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5), 
        axis.text = element_blank(),
        axis.line.x = element_line(size = 1.5),
        axis.line.y = element_line(size = 1.5),
        legend.position = "none")
    
    if (!(length(target) == 0)){
      shaded_df <- data.frame()
      for (nn in 1:length(target)){
        ss <- target[[nn]]
        shaded_df[nn,1:4] <- c(min(ss),max(ss),-Inf,Inf)    
      }
      colnames(shaded_df) <- c("xmin","xmax","ymin","ymax")
      p <- p + geom_rect(data = shaded_df, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
                inherit.aes = FALSE, fill = "gray", colour = NA, alpha = 0.5)
    }
    
    print(p)
    # Save the plot
    ggsave(filename = paste0("../figures/curve_new/",dd, "_", tt,".png"),
           plot = p, width = 5, height = 4)
    
    }

  }
}
```

# Correlation between Gesell_Lang & mean DTI metric (tract-wise level)
```{r}
correlation_results <- data.frame()

    data_avg$Gesell_Lang <- as.numeric(data_avg$Gesell_Lang)
    data_avg$gender <- NA
    data_avg$gender[data_avg$Gender=="M"] <- 1 
    data_avg$gender[data_avg$Gender=="F"] <- 2
    
    #View(data_avg)

i <- 0
for (dd in unique(data_avg$metric)){
  for (tt in unique(data_avg$tractID)){
    for (gg in c("ASD","TD")){
    
    i <- i+1
    data_new <- data_avg %>% filter(metric == dd & tractID == tt &
                                       Dx == gg)
    
    data_new$Gender <- as.factor(data_new$Gender)
    
    model <- lm(avg_value ~ Gesell_Lang + Age + Gender, data = data_new)
    observed_coeff <- summary(model)$coefficients[2, "Estimate"]
    observed_p_value <- summary(model)$coefficients[2, "Pr(>|t|)"]
    
    cor_r_pred <- cor.test(predict(model), data_new$Gesell_Lang)
    cor_r <- cor.test( data_new$Gesell_Lang, data_new$avg_value)
    
    pcor_r <- ppcor::pcor.test(data_new$Gesell_Lang,data_new$avg_value, data_new[,c("Age","gender")])
        
    # Perform permutation test
    #set.seed(123)
    #permuted_coeffs <- numeric()
   # for (j in 1:5000) {
   #   perm_data <- data_new
    #  perm_data$avg_value <- sample(perm_data$avg_value)
   #   permuted_model <- lm(avg_value ~ Gesell_Lang + Age + Gender, data = perm_data)
    #  permuted_coeffs[j] <- summary(permuted_model)$coefficients[2, "Estimate"]
    #  }
   # p_value_perm <- mean(abs(permuted_coeffs) >= abs(observed_coeff))
    
    #correlation_results[i,1:8] <- c(tt,dd,gg,observed_coeff,observed_p_value,
     #                               p_value_perm,cor_r$estimate,pcor_r$estimate)
    
    correlation_results[i,1:8] <- c(tt,dd,gg,observed_coeff,observed_p_value,
                                    cor_r_pred$estimate, cor_r$estimate,pcor_r$estimate)
    }
  }
}

colnames(correlation_results) <- c("tractID","DTI_metric","group",
				   "regression_coefficients",
                                   "P_value","cor_r_value","cor_r_value",
				   "pcor_r_value")

View(correlation_results)

write.csv(correlation_results, "results/correlation_tract_Lang2.csv", 
	  row.names = FALSE)

## test correlation analysis
residuals(data_new$Gesell_Lang, data_new[,c("Age","gender")])



# scatter plots for significant correlation between Gesell_Lang & mean DTI metric 

aa <- read.csv("results/correlation_tract_Lang.csv")

plot_data <- aa %>% filter(perm_p_value < 0.05)

#View(plot_data)

ggplot(data_new[data_new$Dx == "TD", ], aes_string(x = "Gesell_Lang", 
						   y = "avg_value",color = "Dx")) +
    geom_point(size = 3.5) +
   geom_smooth(method = "lm", aes(y = pred),se = FALSE,size=1.8) + 
    #geom_line(aes(y = pred), size=1.8)

cor.test(as.numeric(data_new$Gesell_Lang[data_new$Dx == "TD"]), 
	 as.numeric(data_new$avg_value[data_new$Dx == "TD"]))

for (i in 1:nrow(plot_data)){
	
  data_new <- data_avg %>% filter(tractID == plot_data$tractID[i] &
                                    metric == plot_data$DTI_metric[i])
  
  fl <- formula(avg_value ~ Gesell_Lang + Age + Gender)
  
  models <- data_new %>% split(.$Dx) %>% map(~ lm(fl, data = .x))
  
  data_new <- data_new %>%
    mutate(pred = case_when(
      Dx == "ASD" ~ predict(models[["ASD"]], newdata = .),
      Dx == "TD" ~ predict(models[["TD"]], newdata = .)))
  
  p <-  ggplot(data_new, aes_string(x = "Gesell_Lang", y = "avg_value",color = "Dx")) +
    geom_point(size = 3.5) +
   geom_smooth(method = "lm", se = FALSE,size=1.8) + 
    geom_line(aes(y = pred), size=1.8) + 
    labs(title = paste0(plot_data$DTI_metric[i]," ", plot_data$tractID[i]), 
         x = NULL, y = NULL) +
    theme_classic() +
    theme(axis.ticks.length = unit(0.25, "cm"), 
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5), 
        #axis.text = element_blank(),
        axis.line.x = element_line(size = 2),
        axis.line.y = element_line(size = 2),
        legend.position = "none")+
    scale_color_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC"))
  
  ggsave(filename = paste0("results/Lang_tract_",plot_data$DTI_metric[i],
  			  "_", plot_data$tractID[i],".png"),
         plot = p, width = 6, height = 4)
  
  
}
```

# Correlation between ADOS_Total_ASD & mean DTI metric (tract-wise level)
```{r}
correlation_ados_results <- data.frame()

i <- 0
for (dd in unique(data_avg$metric)){
  for (tt in unique(data_avg$tractID)){
    
    i <- i+1
    data_new <- data_avg %>% filter(metric == dd & tractID == tt &
                                       Dx == "ASD")
    data_new$Gender <- as.factor(data_new$Gender)
    
    model <- lm(avg_value ~ ADOS_Total + Age + Gender, data = data_new)
    observed_coeff <- summary(model)$coefficients[2, "Estimate"]
    observed_p_value <- summary(model)$coefficients[2, "Pr(>|t|)"]
    
    cor_r <- cor.test(predict(model), data_new$ADOS_Total)
        
    # Perform permutation test
    permuted_coeffs <- numeric()
    set.seed(123)
    for (j in 1:5000) {
      perm_data <- data_new
      perm_data$avg_value <- sample(perm_data$avg_value)
      permuted_model <- lm(avg_value ~ ADOS_Total + Age + Gender, data = perm_data)
      permuted_coeffs[j] <- summary(permuted_model)$coefficients[2, "Estimate"]
      }
    p_value_perm <- mean(abs(permuted_coeffs) >= abs(observed_coeff))
    
    correlation_ados_results[i,1:6] <- c(tt,dd,
                                         observed_coeff,
                                         observed_p_value,
                                         p_value_perm,
                                         cor_r$estimate)
    
  }
}

colnames(correlation_ados_results) <- c("tractID","DTI_metric","regression_coefficients",
                                   "P_value","perm_p_value","cor_r_value")

write.csv(correlation_ados_results, "../results/correlation_results_ADOS.csv", row.names = FALSE)

# scatter plots 

aa <- read.csv("results/correlation_results_ADOS.csv")

plot_data <- aa %>% filter(perm_p_value < 0.05)

for (i in 1:nrow(plot_data)){
  data_new1 <- data_avg %>% filter(tractID == plot_data$tractID[i] &
                                    metric == plot_data$DTI_metric[i] &
                                     Dx == "ASD")
  
  fl <- formula(avg_value ~ ADOS_Total + Age + Gender)
  models <- lm(fl, data = data_new1)
  
  data_new2 <- data_new1 %>%
    mutate(pred = predict(models, newdata = data_new1))
  
  p <-  ggplot(data_new2, aes_string(x = "ADOS_Total", y = "avg_value")) +
    #geom_point(size = 3.5, color="black") +
    geom_smooth(method = "lm", se = FALSE, aes(y = pred),color="#e47159",size =1.5) +
    geom_point(size = 5, color="#e47159") +
    geom_smooth(method = "lm", se = FALSE, aes(y = pred),color="black",size =1.8) + 
    labs(title = paste0(plot_data$DTI_metric[i]," ", plot_data$tractID[i]), 
         x = NULL, y = NULL) +
    theme_classic() +
    theme(axis.ticks.length = unit(0.25, "cm"), 
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5), 
        #axis.text = element_blank(),
        axis.line.x = element_line(size = 2),
        axis.line.y = element_line(size = 2),
        legend.position = "none") 
  
  if (plot_data$DTI_metric[i] == "dti_rd") {
  	p <- p + coord_cartesian(xlim = c(5, 25), ylim = c(0.0006, 0.0009))
  } else if (plot_data$DTI_metric[i] == "dti_fa") {
  	p <- p + coord_cartesian(xlim = c(5, 25), ylim = c(0.35, 0.5))
  }
  
  ggsave(filename = paste0("results/",plot_data$DTI_metric[i], "_",
  			 plot_data$tractID[i],"_ADOS.png"),width = 5, height = 4)
  
}
```

# Prepare data for signodes.
```{r}
data_curve <- read.csv("../results/tract_each_node_f_test_results.csv")

# old data
# data_curve <- read.csv("../tables/group_comparison_point_wise/tract_each_node_f_test_results.csv")

avg_dti_signodes <- data.frame() 

for (dd in unique(data_curve$metric)){
  for (tt in unique(data_curve$tractID)){
  # for (tt in unique(data_curve$roi)){
    
     df <- data %>% filter(metric == dd & tractID == tt)
    
    # find consecutive idx 
    data_curve_new <- data_curve %>% dplyr::filter(metric == dd & tractID == tt)
    # data_curve_new <- data_curve %>% dplyr::filter(metric == dd & roi == tt)
    # idx <- which(data_curve_new$p_value_perm < 0.05)
    idx <- which(data_curve_new$p_value < 0.05)
    # idx <- which(data_curve_new$p_value < 0.05)
    # idx <- which(data_curve_new$adjusted_p_value < 0.05)
    idx <- idx - 1
    
    target <- list()
    if (length(idx) >= 3) {
      start <- 1
      for (i in 2:length(idx)) {
        if (idx[i] != idx[i - 1] + 1) {
          if (i - start >= 3) {
            target[[length(target) + 1]] <- idx[start:(i - 1)]
            }
          start <- i 
        }
        }
      if (start <= length(idx) - 2) {
        target[[length(target) + 1]] <- idx[start:length(idx)]
      }
      }
    
    # average sig. nodes
    if (length(target)>0){
      for (j in 1:length(target)){
       df_new <- df %>% filter(nodeID %in% target[[j]])
        
        avg_dti_signodes_sub <- df_new %>%
          group_by(subject_id) %>%
          summarise(avg_value_signodes = mean(value),
                    tractID = first(tractID),
                    metric = first(metric),
                    Gender = first(Gender),
                    Age = first(Age),
                    Dx = first(Dx),
                    Gesell_Lang = first(Gesell_Lang),
                    ADOS_Total = first(ADOS_Total),
                    .groups = 'drop')
        
        avg_dti_signodes_sub <- avg_dti_signodes_sub %>%
          mutate(signodes = paste(range(target[[j]]), collapse = "-") )
        
        avg_dti_signodes <- rbind(avg_dti_signodes, avg_dti_signodes_sub)
      }
    }
    
  }
}
avg_dti_signodes$signodes <- as.character(avg_dti_signodes$signodes )
write.csv(avg_dti_signodes,'results/avg_dti_signodes.csv',row.names = FALSE)


```

# Correlation between Gesell_Lang & sig. nodes (point-wise)
```{r}
 set.seed(123)

avg_dti_signodes <- read.csv("results/avg_dti_signodes.csv")
avg_dti_signodes$gender <- NA
avg_dti_signodes$gender[avg_dti_signodes$Gender == "M"] <- 1
avg_dti_signodes$gender[avg_dti_signodes$Gender == "F"] <- 2

correlation_results <- data.frame()
i <- 0
for (dd in unique(avg_dti_signodes$metric)){
  for (tt in unique(avg_dti_signodes$tractID)){
    
    df <- avg_dti_signodes %>% filter(metric == dd & tractID == tt)
    for (nn in unique(df$signodes)){
      
      df_new <- df %>% filter(signodes == nn)
      
      for (gg in c("ASD","TD")){
        i <- i+1
        data_new <- df_new %>% filter(Dx == gg)
        
        model <- lm(avg_value_signodes ~ Gesell_Lang + Age + Gender, data = data_new)
        observed_coeff <- summary(model)$coefficients[2, "Estimate"]
        observed_p_value <- summary(model)$coefficients[2, "Pr(>|t|)"]
        
        cor_r <- cor.test(predict(model), data_new$Gesell_Lang)
         ppcor <- ppcor::pcor.test(data_new$avg_value_signodes,
         			  data_new$Gesell_Lang,
         			  data_new[,c("Age","gender")])
        
        # Perform permutation test
        permuted_coeffs <- numeric()
        for (j in 1:5000) {
          perm_data <- data_new
          perm_data$avg_value_signodes <- sample(perm_data$avg_value_signodes)
          permuted_model <- lm(avg_value_signodes ~ Gesell_Lang + Age + Gender,
                               data = perm_data)
          permuted_coeffs[j] <- summary(permuted_model)$coefficients[2, "Estimate"]
          }
        p_value_perm <- mean(abs(permuted_coeffs) >= abs(observed_coeff))
        
        correlation_results[i,1:9] <- c(tt,dd,gg,nn,
                                        observed_coeff,
                                        observed_p_value,
                                        p_value_perm,
                                        cor_r$estimate, ppcor$estimate)
      }
    }
  }
}

colnames(correlation_results) <- c("tractID","DTI_metric","Dx","signodes",
                                   "regression_coefficients",
                                   "p_value","perm_p_value","cor_r_value",
				   "PPcor_r_value")

View(correlation_results)

write.csv(correlation_results, "results/correlation_signodes_Lang.csv",
	  row.names = FALSE)

# scatter plots for significant correlation between Gesell_Lang & sig. nodes (point-wise)

aa <- read.csv("results/correlation_signodes_Lang.csv")

data_scatter <- aa %>% filter( p_value < 0.05)
sig_nodes <- c(17:21)

df_avg <- avg_dti_signodes

for (dd in unique(data_scatter$DTI_metric)){ 
  for (tt in unique(data_scatter$tractID)){

    tmp <- data_scatter %>% filter(DTI_metric == dd & tractID == tt)
    sig_nodes <- as.numeric(strsplit(tmp$signodes,"-")[[1]])
    
    df <- df_avg %>% filter(metric == dd & tractID == tt &
                                 signodes %in% tmp$signodes)
  
    
    # predict
    formula <- formula(avg_value_signodes ~ Gesell_Lang + Age + Gender)
    model_ASD <- lm(formula, data = df[df$Dx == "ASD",])
    model_TD <- lm(formula, data = df[df$Dx == "TD",])
    pred_ASD <- predict(model_ASD, newdata = df[df$Dx == "ASD",])
    pred_TD <- predict(model_TD, newdata = df[df$Dx == "TD",])
    df$pred <- NA
    df$pred[df$Dx == "TD"] <- pred_TD
    df$pred[df$Dx == "ASD"] <- pred_ASD
    
    p <- ggplot(df, aes(x = Gesell_Lang, y = avg_value_signodes,color = Dx)) +
      geom_point(size = 3.5) +
      stat_smooth(method = "lm", se = FALSE, 
      	    aes(y = pred),size=1.8) +
      labs(title = paste0(dd," ", tt), 
           x = NULL, y = NULL) +
       theme_classic() +
      theme(axis.ticks.length = unit(0.25, "cm"), 
        #axis.text = element_blank(),
        axis.line.x = element_line(size = 2),
        axis.line.y = element_line(size = 2),
        legend.position = "none")+
      scale_color_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC"))
    
    # Save the plot
    ggsave(paste0("results/",dd,"_",tt,"_Gesell_Lang_signodes.png"), 
           plot = p, width = 6, height = 4)
    
    
  }
}

library(ggplot2)
library(dplyr)

# Fit the model
model <- lm(avg_value ~ Gesell_Lang + Age + Gender, data = data_new)

# Create a new data frame for predictions across a range of Gesell_Lang for each group
prediction_data <- data_new %>%
  group_by(Dx) %>%
  summarize(Age = mean(Age),       # Use average Age for predictions
            Gender = "Male",       # Set Gender to one category; adjust as necessary
            Gesell_Lang = seq(min(Gesell_Lang), max(Gesell_Lang), length.out = 100)) %>%
  ungroup()

# Add predicted values to the prediction_data
prediction_data <- prediction_data %>%
  mutate(predicted_values = predict(model, newdata = prediction_data))

# Plot the data and the fitted lines
ggplot(data_new, aes(x = Gesell_Lang, y = avg_value, color = Dx)) +
  geom_point(size = 3) +  # Scatter plot points for original data
  geom_line(data = prediction_data, aes(y = predicted_values), size = 1.2) + # Fitted line from predictions
  labs(title = "Fitted Regression Lines for ASD and TD Groups",
       x = "Gesell Language Score",
       y = "Average Value") +
  theme_minimal() +
  scale_color_manual(values = c("ASD" = "#e47159", "TD" = "#6699CC"))

```

# Correlation between ADOS & sig. nodes (point-wise) 
```{r}
set.seed(123)
avg_dti_signodes <- read.csv("results/avg_dti_signodes.csv")
correlation_results <- data.frame()

i <- 0
for (dd in unique(avg_dti_signodes$metric)){
  for (tt in unique(avg_dti_signodes$tractID)){
    
    df <- avg_dti_signodes %>% filter(metric == dd & tractID == tt & Dx == "ASD")
    for (nn in unique(df$signodes)){
      i <- i +1
      df_new <- df_avg %>% filter(signodes == nn)
      
      model <- lm(avg_value_signodes ~ ADOS_Total + Age + Gender, data = df_new)
      observed_coeff <- summary(model)$coefficients[2, "Estimate"]
      observed_p_value <- summary(model)$coefficients[2, "Pr(>|t|)"]
      cor_r <- cor.test(predict(model), df_new$ADOS_Total)
      
      # Perform permutation test
      permuted_coeffs <- numeric()
      for (j in 1:5000) {
        perm_data <- df_new
        perm_data$avg_value_signodes <- sample(perm_data$avg_value_signodes)
        permuted_model <- lm(avg_value_signodes ~ Gesell_Lang + Age + Gender, data = perm_data)
        permuted_coeffs[j] <- summary(permuted_model)$coefficients[2, "Estimate"]
       }
      
      p_value_perm <- mean(abs(permuted_coeffs) >= abs(observed_coeff))
      
      correlation_results[i,1:7] <- c(tt,dd,nn,
                                      observed_coeff,
                                      observed_p_value,
                                      p_value_perm,
                                      cor_r$estimate)
      }
  }
}

colnames(correlation_results) <- c("tractID","DTI_metric","signodes",
                                   "regression_coefficients",
                                   "p_value","perm_p_value","cor_r_value")

View(correlation_results)

write.csv(correlation_results, "results/correlation_results_signodes_ADOS.csv", 
	  row.names = FALSE)

## scatter plots between sig. nodes & ADOS (point-wise)

aa <- read.csv("results/correlation_results_signodes_ADOS.csv")

data_scatter <- aa %>% filter(p_value < 0.05)


for (i in 1:nrow(plot_data)){
  tmp <- data_scatter %>% filter(DTI_metric == data_scatter$DTI_metric[i]  &
                                   tractID == data_scatter$tractID[i])
  #sig_nodes <- as.numeric(strsplit(tmp$signodes,"-")[[1]])
  
  df <- df_avg %>% filter(metric == data_scatter$DTI_metric[i] & 
                               tractID == data_scatter$tractID[i] &
                               signodes %in% tmp$signodes & 
                               Dx == "ASD")
    
    # predict
    ff <- formula(avg_value_signodes ~ ADOS_Total + Age + Gender)
    model <- lm(ff, data = df)
    df$pred <- predict(model, newdata = df)
    
    title <- paste0(data_scatter$DTI_metric[i]," ", 
                          data_scatter$tractID[i])
    
    p <- ggplot(df, aes(x = ADOS_Total, y = avg_value_signodes)) +
      geom_point(size = 5, color="#e47159") +
      geom_smooth(method = "lm", se = FALSE, aes(y = pred),
                  color="black",size = 1.8) +
      labs(title = title, 
           x = NULL, y = NULL) +
      theme_classic() +
      theme(axis.ticks.length = unit(0.25, "cm"), 
        plot.title = element_text(size = 20, face = "bold", hjust = 0.5), 
        #axis.text = element_blank(),
        axis.line.x = element_line(size = 2),
        axis.line.y = element_line(size = 2),
        legend.position = "none")
    
    
  if (plot_data$DTI_metric[i] == "dti_rd") {
  	p <- p + coord_cartesian(xlim = c(5, 25), ylim = c(0.0006, 0.0009))
  } else if (plot_data$DTI_metric[i] == "dti_fa") {
  	p <- p + coord_cartesian(xlim = c(5, 25), ylim = c(0.35, 0.5))
  }
}
    
    # Save the plot
    ggsave(paste0("results/",title,"_ADOS_signodes.png"), width = 5, height = 4)

    

```


